---
title: "Titanic Survivor Data Exploration"
author: "Saixiong Han, En Liu, Liping Li"
date: "October 20, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Train Data Processing &Description
>By Liping Li

###-Import and Overview

```{r echo=TRUE,warning=FALSE,tidy=TRUE}
library(data.table)
train=fread("C:/data/train-titanic.csv", sep=",", header=TRUE, stringsAsFactors=TRUE)
setkey(train)
summary(train)
```



###-Deal with Age's NAs

Age has 177 missing values. Cabin has 687 NAs and Embarked has 2. We'll can put Cabin and Embarked aside. Age is an important demographic variable, so we can try to fill with predictions.

```{r echo=TRUE,warning=FALSE,tidy=TRUE}
train$Embarked=as.character(train$Embarked) #change data type that can add NA
train[Embarked=='', Embarked:=NA] #set '' in Embarked as missing value
```

```{r echo=TRUE,message=FALSE, warning=FALSE,tidy=TRUE,fig.height=3}
# Tried regression model, R-square is not ideal. So take the tree method.
library(rpart)
library(ggplot2)
library(gridExtra)
age_fit2 = rpart(Age ~ Pclass + Sex + SibSp + Parch
                +Fare+Embarked, 
                data = train[!is.na(Age), ],
                na.action = na.omit, method = 'anova',cp=0.01)
page1=qplot(train$Age,geom='histogram',fill=I('turquoise3'),xlab='Age-raw', 
           ylab='Numbers of People')
train[is.na(Age), Age := predict(age_fit2, newdata=train[is.na(Age)]) ]
page2=qplot(train$Age,geom='histogram',fill=I('turquoise3'),xlab='Age-adjusted', 
            ylab='Numbers of People')
grid.arrange(page1,page2, ncol=2)
```

The two distribution of age seem alike. That's what we want.


###-List single variable distribution except Name, Ticket and Cabin

Now, most of variables are factors and numbers. It's like that you finally pick all the vegetables and meat for dinner. Cannot wait to have a raw view.

```{r echo=TRUE,message=FALSE, warning=FALSE,tidy=TRUE, fig.height=9,fig.width=10}
psur=qplot(train$Survived,geom='bar', ylab='Numbers of People',fill=I('turquoise3'), 
           xlab="0=Not survived   1=Survived" )
ppclass=qplot(train$Pclass,geom='bar',fill=I('turquoise3'),xlab='Social-Economic Status', 
      ylab='Numbers of People')
psex=qplot(train$Sex,geom='bar',fill=I('turquoise3'), xlab='Gender',
      ylab='Numbers of People')
pss=qplot(train$SibSp,geom='bar',xlab='People with x Siblings and Spouse', 
      ylab='Numbers of People',fill=I('turquoise3'))
ppc=qplot(train$Parch,geom='bar',xlab='People with x Parents and Children', 
      ylab='Numbers of People',fill=I('turquoise3'))
pf=qplot(train$Fare, geom='histogram',fill=I('turquoise3'),xlab='Price of Ticket', 
      ylab='Numbers of People')
pe=qplot(train$Embarked, xlab='Embarked Port', fill=I('turquoise3'),
      ylab='Numbers of People',geom='bar')
grid.arrange(psur, ppclass,psex,page2,pss,ppc,pf,pe, ncol=3)
```

Most people are in their golden ages(20~40). Most people are men. Most people are travel alone. Most people from 3rd class. Most people from Southampton. And most of them lost their lives in that adventure.


###-Interesting Thing about Name

Name has many interesting title. Let's grasp these info.
```{r echo=TRUE,warning=FALSE,tidy=TRUE}
train$Title=gsub('(.*,)|(\\..*)','',train$Name)
train$Title=as.factor(train$Title)
train$Survived=as.character(train$Survived)
ggplot(train,aes(x=Title, fill=Survived))+geom_bar()+coord_flip()+scale_fill_brewer(palette = "BuGn")+ylab('')
```

The survived "Miss"" is **no** more than married women "Mrs"" in terms of ration. I've been heard that people in Titanic give survivor opportunities to unmarried women. But according to this chart, it seems not ture. We'll dig into it later.

Another phenomenon attracts me is that many people have two names, like "Cumings, Mrs. John Bradley (Florence Briggs Thayer)". Did they change name before this tragedy or after? We don't know. Just have a look.

```{r echo=TRUE,warning=FALSE,tidy=TRUE,fig.height=3,fig.width=5}
train$NewName=grepl('(', train$Name,fixed=TRUE)
train$Survived=as.character(train$Survived)
ggplot(train,aes(x=NewName, fill=Survived))+geom_bar()+scale_fill_brewer(palette = "BuGn")+ylab('')
```

The ration looks high for people survived to have a second name. 


###-A Quick Look of Correlation and Importance
```{r echo=TRUE,message=FALSE, warning=FALSE,tidy=TRUE}
library(GGally) #Only shows numeric variables
train[Sex=='male',nSex:=1]
train[Sex=='female',nSex:=2]
train[Embarked=='C',Port:=1]
train[Embarked=='Q',Port:=2]
train[Embarked=='S',Port:=3]
train$Survived=as.numeric(train$Survived)
cor=ggcorr(train,label=TRUE, label_alpha=TRUE, name='Correlation', low = "turquoise3", mid = "white", high = "orangered3")

library(randomForest)
rf = randomForest(Survived ~ Pclass+Age+nSex+Parch+SibSp+Fare+Port, data=train,
                   na.action = na.omit, ntree=101, importance=TRUE)
imp = importance(rf, type=1)
featureImportance = data.frame(Feature=row.names(imp), Importance=imp[,1])
ggplot(featureImportance, aes(x=reorder(Feature, Importance), y=Importance)) +
  geom_bar(stat="identity", fill=I("turquoise3")) +
  coord_flip() +  xlab('')+ylab("Importance on Survive")
```

**Women**, people of **young age** and people of **higher class** are more likely to survive.


###-X with Survived

Take a close look of every independent variable influence on survive.
This time add SibSp and Parch as family size
```{r echo=TRUE,message=FALSE, warning=FALSE,tidy=TRUE,fig.height=6,fig.width=10}
train$Survived=as.character(train$Survived)
SurPc=ggplot(train,aes(x=Pclass, fill=Survived))+geom_bar()+ scale_fill_brewer(palette = "BuGn")+ylab('')
Surage=ggplot(train,aes(x=Age, fill=Survived))+geom_histogram()+ scale_fill_brewer(palette = "BuGn")
SurFam=ggplot(train,aes(x=SibSp+Parch, fill=Survived))+geom_histogram()+scale_fill_brewer(palette = "BuGn")+ylab('')
SurFare=ggplot(train,aes(x=Fare, fill=Survived))+geom_histogram()+ scale_fill_brewer(palette = "BuGn")+ylab('')
SurSex=ggplot(train,aes(x=Sex, fill=Survived))+geom_bar()+ scale_fill_brewer(palette = "BuGn")+ylab('')
SurEmbarked=ggplot(train,aes(x=Embarked, fill=Survived))+geom_bar()+ scale_fill_brewer(palette = "BuGn")+ylab('')
grid.arrange(SurPc,Surage,SurFam,SurFare,SurSex,SurEmbarked, ncol=3) 
```

From the diagram, we have a guess: people who are women, young but not too young, have 1-2 companies have more chance to survive.



##2. Calculate Single Factor Importance in Female Surviving
>Saixiong Han

### -Pclass

```{r Pclass, echo=FALSE}
table(train$Sex,train$Pclass,train$Survived)
```

The female survived in the Titanic most from first and second class, 
the number of female in the first and second class died in Titanic is less than 10. 
This is a very important factor.

### -Parch

```{r Parch, echo=FALSE}
table(train$Sex,train$Parch,train$Survived)
```
Women without any children or parents are more likely to survive 
While the more parch they have, the less probability they can survive. 

### -SibSp

```{r SibSp, echo=FALSE}
table(train$Sex,train$SibSp,train$Survived)
```
Women without any siblings or spouses are more likely to survive.
While the more siblings they have, the less probability they can survive. 

### -Title

```{r title, echo=FALSE}
table(train$Sex,train$Title,train$Survived)
```
The survived unmarried women is more than married women, which is accordance with the result about SibSp. Generally, the number of women survived is more than the women died.

### -Data visualization

####**Pclass,Survived,Age and Fare Scatter plot**

```{r scatter1, echo=FALSE}
train$Pclass=as.character(train$Pclass) #To make sure change safety
train$Pclass=as.factor(train$Pclass)
p<-ggplot(data=train,mapping = aes(x=Age,y=Fare,shape=Pclass,color=Survived))
p+geom_point()
```

This scatter plot shows there is no linear relation between age and fare. 
From the color scatter we can see, blue points are mostly on the top wihle the red points are maily on the bottom. 
That means the higher fare is, the more chances for surviving. It also can be shown in the shape of different classes. The first and second class get more chances for surviving compared to the third class.

####**Sex,Survived,Age and Fare Sctter plot**

```{r scatter2, echo=FALSE}
p<-ggplot(data=train,mapping = aes(x=Age,y=Fare,shape=Sex,color=Survived))
p+geom_point()+scale_fill_brewer(palette = "Paired")
```

the blue round points is more than blue squre points. 
That shows females have more chances of surviving compared to males. 
In addition, the higher fare of females bought, the higher chances for surviving.

####**Fare,Sex and Survived(box-plot)**

```{r boxplot1, echo=FALSE}
ggplot(train)+geom_boxplot(aes(x=Sex,y=Fare,fill=Survived))
```
The lower whisker and lower quartile for all the four box are the same.
the survived female have higher median and upper quartier than the died female.
That shows female with higher fare or higher pclass have more chance for surviving.
it's the same with male's surviving, but the amount of surviving men is smaller than women.
The distribution of survived female is very sparse while the distribution of died male is very intensive.

####**Fare,Pclass and Survived(box-plot)**

```{r boxplot2, echo=FALSE}
ggplot(train)+geom_boxplot(aes(x=Pclass,y=Fare,fill=Survived))
```

The charts shows that the fare in the first class has larger range than the second and third class.
The median of survived people in first class is equal to the upper quartier of dead people.
Generally, people in the first class with higher fare have more chance to survive.
The upper whisker of second and third class are almost equal to the lower quartier of first class.
The fare distribution of second and third class are more intensive than the first class.


####**Age,Pclass and Survived(violin_plot)**

```{r violin1, echo=FALSE}
p<-ggplot(train,aes(factor(Pclass),Age))
p+geom_violin(aes(fill=Survived))
```

####**Age,Sex and Survived(violin_plot)**

```{r violin2, echo=FALSE}
p<-ggplot(train,aes(factor(Sex),Age))
p+geom_violin(aes(fill=Survived))
```



##3.Analytics of factors
>En Liu

Let's do some investigation at the influence of age, pclass, and sex to the survival. In order to show the result vividly, we use ggplot2 to draw the facet_grid diagram.

```{r facet1}
result1 = ggplot(data = train, aes(x = Age, y = Sex, color = factor(Survived)))+geom_point()
result = result1+facet_grid(.~Pclass)
result
```
In this diagram, we can see that age, pclas and sex are all important factors. For those who were in the first class and who were female, there was a bigger chance to survive. We can imagine the scene back then: Titanic was falling, few lifeboats were prepared and young, well-classed ladies and madams were rescued first.For those who landed lifeboats and left, the chance of survival was the biggest.But for those females who were in the third class and the males, things were not so happy. Lifeboats were limited, they must find their own way to make lives. The possiblities of their living can be very small.

Also, we can discuss the relationship between survival and family size. There are no factors called "family size", but we can conclude from siblings and parch. According to our experience, those who were with babies and  old people might have a bigger chance to survive. We can know if we are right from the following diagram.

```{r facet2}
family = train$SibSp+train$Parch
result2 = ggplot(data = train, aes(x = Age, y = family, color = factor(Survived)))+geom_point()
result = result2+facet_grid(.~Sex)
result
```

Then we can see that our assumption is true. First,females with 1 or 5,6 members in a family got the biggest chance of survival, while those who had only 1 or 2 family members did not. We can imagine that they might honorly give the chance of living to those who need most, and choose to die with their husbands or fathers. Males does not have a significant feature from the diagram. 
